### 2022-09-23 chapters/benchmarking/benchmarking.tex
Proofreading titles

\newpage
\chapter{A Benchmarking Protocol to Measure Software Energy Consumption}\label{chapter:benchmarking}

\import{\currfiledir}{draft}
\import{\currfiledir}{reproducibility}
\import{\currfiledir}{experiment}
\import{\currfiledir}{conclusion}

>>>>

\chapter{Benchmarking Protocol to Measure Software Energy Consumption}\label{chapter:benchmarking}
% \section{Introduction \note{missing} }
This chapter covers ways to overcome empirical analysis challenges in energy consumption studies.
First, we go through the three components of a successful benchmark when performing energy-related experiments.
Section~\ref{sec:benchmarking_reproducibility} focuses on the "reproducibility" challenge---to deliver reproducible experiments without interfering with the energy measurement---while Section~\ref{sec:taming-the-energy-variation} discusses accuracy of software energy consumption.

\import{\currfiledir}{draft}
\import{\currfiledir}{experiment}
\import{\currfiledir}{conclusion}

### 2022-09-23 chapters/optimization/java-dir/java.tex
refactoring
corrections : end part

\subsection{characteristics of JVM}
\begin{figure}%[ht]
    \centering
    \includegraphics[width=\linewidth]{imgs/Jvm_architecture}
    \caption{JVM architecture}
    \label{fig:JVM_architecture}
\end{figure}

@@

\subsection{Goal}
In this section, we will investigate the following research questions:
\begin{compactenum}[\indent\bf RQ\,1:]
    \item \emph{What is the impact of existing JVM distributions on the energy consumption of Java-based software services?}
    \item \emph{What are the relevant JVM settings that can reduce the energy consumption of a given software service?}
\end{compactenum}

@@

\section{Conclusion}
The results of our investigations show that many JVMs share energy efficiencies and can be grouped into 3 classes: \textsc{HotSpot}, J9, and \textsc{GraalVM}.
The 3 selected JVM classes can however report a different energy efficiency for different software and/or workloads, sometimes by a large margin.
While we did not observe a unique champion when it comes to energy consumption, \textsc{GraalVM} reported the best energy efficiency for a majority of benchmarks.
Nonetheless, each JVM can achieve the best or the worst depending on the hosted application.
One cause can be thread management strategies, as observed with J9 when advantageously running \textsf{Avrora}.
Moreover, some JVM settings can cause energy consumption variations.
Our experiments showed that the default JIT compiler of the JVM is often near-optimal, in at least ~80\% of our experiments.
The default GC, however, was outperforming alternative strategies in half of our experiments, with some large gains observed when using some alternative GC depending on the application characteristics.

Our main conclusions and guidelines can be thus summarized as:
\begin{inparaenum}[\em i)]
    \item testing software on the 3 classes of JVM and identifying the one that consumes the least is a good practice, especially for multi-threading purposes,
    \item while the JVM default JIT give often good energy consumption results, some settings may improve the energy consumption and could be tested,
    \item the choice of the GC may lead to a large impact on the energy consumption in many situations, thus encouraging a careful tuning of this parameter before deployment.
\end{inparaenum}
To ease the integration of the above guidelines, we propose a tool, named \textsf{J-Referral}, to recommend the most energy-efficient JVM distribution and configuration among more than a hundred considered possibilities.
It establishes a full report on the energy consumption of both CPU and DRAM components for each JVM distribution and/or configuration to help the user to choose the one with the least consumption for Java software.
% TODO : add or change the graphes of the use case 
% TODO : add the description and a diagrame for the experimental protocol

>>>>

\subsection{Characteristics of JVM}
\begin{figure}%[ht]
    \centering
    \includegraphics[width=\linewidth]{imgs/Jvm_architecture}
    \caption{JVM architecture}
    \label{fig:JVM_architecture}
\end{figure}

@@

\subsection{Research questions}
In this section, we will investigate the following research questions:
\begin{compactenum}[\indent\bf RQ\,1:]
    \item \emph{What is the impact of existing JVM distributions on the energy consumption of Java-based software services?}
    \item \emph{What are the relevant JVM settings that can reduce the energy consumption of a given software service?}
\end{compactenum}

@@

\clearpage
\section{Summary}

This chapter describes an empirical investigation of contrast in energy consumption that some of the most well-known and widely supported JVM platforms can exhibit as well as the important factors that can influence this energy consumption.
During our experiments, we investigated 12 well-known and diverse-purpose Java benchmarks, as well as 52 JVMs, including many versions of 11 different distributions. 
Our findings revealed that various JVMs share energy efficiency and may be classified into three major classes: \textsc{HotSpot}, \textsc{J9}, and \textsc{GraalVM}.
However, the three JVM classes can show substantially different energy efficiency depending on the workloads and/or application.
While there was no clear champion in terms of energy consumption,\textsc{GraalVM} recorded the highest energy efficiency for the most of benchmarks. Nevertheless, depending on the workload, each JVM can achieve higher or worse efficiency.
One explanation could be thread management policy, as seen with \textsc{J9} when running \textsf{Avrora}.

Additionally, numerous JVM settings can cause changes in energy consumption. Our experiments revealed that the JVM's default JIT compiler is frequently near-optimal in at least 80\% of the experiments.

On the other hand , the default GC outperformed the other strategies in just half of our experiments, with some significant benefits reported when employing different GCs based on the application characteristics. Our primary observations and recommendations can be stated as follows:
\begin{inparaenum}[\em i)]
    \item testing software on the 3 classes of JVM and identifying the one that consumes the least is a good practice, especially for multi-threading purposes,
    \item while the JVM default JIT give often good energy consumption results, some settings may improve the energy consumption and could be tested,
    \item the choice of the GC may lead to a large impact on the energy consumption in many situations, thus encouraging a careful tuning of this parameter before deployment.
\end{inparaenum}

To make the above principles easier to implement, we propose \emph{J-Referral}, a tool that recommends the best energy-efficient JVM distribution and configuration from among more than a hundred options.
It generates a comprehensive report on the energy consumption of both CPU and DRAM components for each JVM distribution and/or configuration in order to assist the user in selecting the one with the lowest consumption for Java applications.

### 2022-09-23 chapters/optimization/python-dir/python_interperters.tex

To help practitioners to choose the best implementation for their use case, we have created a chart that describes the behavior of each implementation toward a specific aspect of programming.
Figure~\ref{fig:tommi_all} introduces a radar plot for all the implementations that have been used in this experimentation.

@@

As one can notice in Figure~\ref{fig:tommi_all}, there is no evolution between CPython2 and CPython3.
Moreover, IntelPython, and ActivePython all behave similarly to the reference implementation.
Therefore, one can conclude that the work done on those interpreters is primarily to improve a specific purpose and not the core interpreter.
ActivePython states that their version is focused on security and prepackaged libraries, which explains why it is slower than other versions due to the addition of this security support.\footnote{\url{https://www.activestate.com/solutions/why-activestate/}}
As for IntelPython, it is designed for machine learning.
Unfortunately, the \textsc{Tommti} benchmark is geared for general-purpose programming.
Despite the fact that Nuitka is a compiler, there was no discernible difference in energy consumption.
It was even more similar to the reference Python implementation than other interpreters.
However, if we look at the Nuitka techniques, we can see that they simply encapsulate the Python code with an interpreter into a single executable.
Finally, Shedskin reports on the best energy consumption pattern when it comes to arithmetic operations.
One can conclude it is due to the fact of the native type of the variables, unlike in the JIT, where they are treated as objects in the beginning.

@@

\subsection{Conclusion}
One can conclude that the choice of Python interpreter has a significant impact on the programs' energy consumption.
While most of the candidates used Python2, we observed that different strategies that were implemented within each candidate have a significant impact on the energy consumption. 
The major improvement of the energy was related to the use of JIT, which is a technique that compiles the code into machine code at runtime. 
The effect of the compilation of the python code into machine code was also confirmed with the use of translators such as shedskin.

This investigation is made more intriguing by the absence of a universal solution.
The primary downside is the incompatibility of some of these solutions, which causes us to make concessions when we need a generic answer.

>>>>

%  Switch this to conclusion 
To help practitioners to choose the best implementation for their use case, we have created a chart that describes the behavior of each implementation toward a specific aspect of programming.
Figure~\ref{fig:tommi_all} introduces a radar plot for all the implementations that have been used in this experimentation.

@@

As one can notice in Figure~\ref{fig:tommi_all}, there is no evolution between CPython2 and CPython3.
Moreover, IntelPython, and ActivePython all behave similarly to the reference implementation.
Therefore, one can conclude that the work done on those interpreters is primarily to improve a specific purpose and not the core interpreter.
ActivePython states that their version is focused on security and prepackaged libraries, which explains why it is slower than other versions due to the addition of this security support.\footnote{\url{https://www.activestate.com/solutions/why-activestate/}}
As for IntelPython, it is designed for machine learning.
Unfortunately, the \textsc{Tommti} benchmark is geared for general-purpose programming.
Despite the fact that Nuitxka is a compiler, there was no discernible difference in energy consumption.
It was even more similar to the reference Python implementation than other interpreters.
However, if we look at the Nuitka techniques, we can see that they simply encapsulate the Python code with an interpreter into a single executable.
Finally, Shedskin reports on the best energy consumption pattern when it comes to arithmetic operations.
One can conclude it is due to the fact of the native type of the variables, unlike in the JIT, where they are treated as objects in the beginning.

@@

\subsection{Conclusion}
One can conclude that the choice of Python interpreter has a significant impact on the programs' energy consumption.
While most of the candidates used Python2, we observed that different strategies that were implemented within each candidate have a significant impact on the energy consumption. 
The major improvement of the energy was related to the use of JIT, which is a technique that compiles the code into machine code at runtime. 
The effect of the compilation of the python code into machine code was also confirmed with the use of translators such as shedskin. 
This investigation is made more intriguing by the absence of a universal solution.
The primary downside is the incompatibility of some of these solutions, which causes us to make concessions when we need a generic answer.
we did a classification and clusturing of different python implementations based on their energy consumption.

### 2022-09-23 chapters/programming_languages/rpc/rpc.tex
refactoring

\subsection{Results \& Findings}
\paragraph{\textsc{RQ}~1: How do RPC implementations consume with regard to the size of the incoming request?}
The purpose of this question is to study the energy consumption of the RPC server when transferring large objects.
To do so, we send $80,000$ requests to the server whose size scales from 10 bytes up to 10 Megabytes, resulting in $10,000$ requests per size per server.
To eliminate extra factors, we let the server handle the rate at which it can answer each request.
However, we put a 20 seconds timeout deadline for each request.
Therefore, our boundary condition is only the number of requests received by the server.
For this experiment, we investigate 4 observable variables:
\begin{enumerate}
    \item the \textsf{average power consumption} during the scenario indicates the overall behavior of the server when working for long durations,
    \item the \textsf{tail latency} for the 99th percentile, which indicates how efficient is the server,
    \item the \textsf{average number of requests per second}, which indicates the average number of clients that the server can handle,
    \item the \textsf{average energy cost} of a single request: unlike the first indicator, this one shows how green is the implementation considering performance.
\end{enumerate}

@@

\subsection{Threads to Validity}

\subsection{Conclusion}

>>>>

\subsection{Results \& Findings}
\paragraph{\textsc{RQ}~1: How do RPC implementations consume with regard to the size of the incoming request?}

The purpose of this question is to study the energy consumption of the RPC server when transferring large objects.
To do so, we send $80,000$ requests to the server whose size scales from 10 bytes up to 10 Megabytes, resulting in $10,000$ requests per size per server.
To eliminate extra factors, we let the server handle the rate at which it can answer each request.
However, we put a 20 seconds timeout deadline for each request.
Therefore, our boundary condition is only the number of requests received by the server.
For this experiment, we investigate 4 observable variables:
\begin{enumerate}
    \item the \textsf{average power consumption} during the scenario indicates the overall behavior of the server when working for long durations,
    \item the \textsf{tail latency} for the 99th percentile, which indicates how efficient is the server,
    \item the \textsf{average number of requests per second}, which indicates the average number of clients that the server can handle,
    \item the \textsf{average energy cost} of a single request: unlike the first indicator, this one shows how green is the implementation considering performance.
\end{enumerate}

@@

% \subsection{Threads to Validity \note{missing}}

\subsection{Conclusion \note{missing}}
